---
title: "WHO/UN Data Life Expectancy"
format: html
editor: visual
---

# Life Expectancy Data Modeling Project Proposal:

### Team Members:

#### Poojitha Balamurugan, Minling Zhou, John Coogan, Ofosu Osei

### Data Source:

Data compiled from WHO and United Nations website by Kumar Rajarshi, Deeksha Russel, and Duan Wang (proposed augmentation from UN economic data).

### Overview:

This data set investigates the intersection between a myriad of societal factors as an influence on life expectancy. Assessing common health data from the World Health Organization by country for variables such as infant mortality, adult mortality, alcohol consumption, hepatitis B immunization coverage, and BMI and introducing economic factors such as expenditure on health as a percentage of GDP, this data attempts to include non-standard factors in its analysis of human health as measured by life expectancy.

### Future Vectors:

This data set takes a comprehensive approach to inclusion of health factors which influence life expectancy. It does not have a thorough inclusion of economic factors which we believe are intrinsically linked to life expectancy. We would like to add GDP per-capita, Gini Index, and Household final consumption expenditure as additional variables (All sourced from the UN data explorer). We believe this will be a more comprehensive approach to assessing factors for life expectancy.

It is our assumption that global health is more than a health-system issue and that investigating what the largest health factors are for life expectancy omits key, and dominant, economic factors. This data set will allow us to explore questions of global health in the context of national economic growth, wealth distribution, and consumer expenditure.

### Research Questions:

1.  Which national economic factors have the largest effect on life expectancy (GDP, Expenditure on health as a percentage of GDP, Wealth inequality, or base household expenditures)
2.  Do national economic factors always directly correlate to better health outcomes (are there rich countries with poorer outcomes/ are there poorer countries whose health systems outperform rich countries). This is an interesting question because it can identify pre-existing healthcare models that are perhaps most efficient.
3.  Categorize the relationship between economic growth and health outcomes (expected nonlinear relationship)

```{r}
library(readr)
library(tidyverse)
life_exp <- read_csv("Life Expectancy Data.csv",show_col_types = FALSE)
```

```{r}
glimpse(life_exp)
```

```{r}
variable.names(life_exp)
```
